
```{r, child=here::here('vignettes', 'includes', '_header.Rmd')}
# Imports custom latex functions for equations
```

# logitr

**logitr** estimates multinomial (MNL) and mixed logit (MXL) models in R. Models can be estimated using "Preference" space or "Willingness-to-pay" (WTP) space utility parameterizations. The current version includes support for:

- Homogeneous multinomial logit (MNL) models
- Heterogeneous mixed logit (MXL) models (with normal and log-normal parameter distributions).
- Preference space utility parameterization.
- WTP space utility parameterization.
- An optional multistart optimization that uses different random starting points in each iteration (useful for non-convex problems like MXL models or models with WTP space parameterizations).
- A simulation function for computing the expected shares of a set of alternatives using an estimated model.

MXL models assume uncorrelated heterogeneity covariances and are estimated using maximum simulated likelihood based on the algorithms in [Kenneth Train's](http://eml.berkeley.edu/~train/) book [*Discrete Choice Methods with Simulation, 2nd Edition (New York: Cambridge University Press, 2009)*](http://eml.berkeley.edu/books/choice2.html).

## Installation

The current version is not yet on CRAN, but you can install it from Github using the **devtools** library:

```
devtools::install_github('jhelvy/logitr')
```

## Required Libraries

**logitr** requires the [**nloptr**](https://cran.r-project.org/web/packages/nloptr/index.html) library. This is because `nloptr()` allows for both the objective and gradient functions to be computed in a single function. This speeds up computation time considerably because both the objective and gradient functions require many of the same calculations (e.g. computing probabilities).

# Estimating models

The `logitr()` function estimates multinomial logit (MNL) and mixed logit (MXL) models using "Preference" space or "Willingness-to-pay (WTP)" space [utility parameterizations](articles/utility_models.html). The basic usage for estimating models is as follows:

```{r, eval=FALSE}
model <- logitr(
  data,
  choiceName,
  obsIDName,
  parNames,
  priceName = NULL,
  randPars = NULL,
  randPrice = NULL,
  modelSpace = "pref",
  weightsName = NULL,
  options = list()
)
```

## Data format

The `data` argument is set to the name of the `data.frame` storing the choice observations. The `data.frame` must be arranged such that each row is an alternative from a choice observation. The choice observations do not have to be symmetric (i.e. each choice observation could have a different number of alternatives). The data must include columns for each of the following arguments in the `logitr()` function:

- `choiceName`: A dummy variable that identifies which alternative was chosen (`1` = chosen, `0` = not chosen).
- `obsIDName`: A sequence of numbers that identifies each unique choice occasion. For example, if the first three choice occasions had 2 alternatives each, then the first 9 rows of the `obsID` variable would be `1, 1, 2, 2, 3, 3`.
- `parNames`: The names of the variables that will be used as model covariates. For WTP space models, do **not** include the price variable in `parNames` - this is provided separately with the `priceName` argument.

## Preference space models

Unlike similar packages like [**mlogit**](https://cran.r-project.org/web/packages/mlogit/index.html), **logitr** does not support the formula type input that uses a symbolic description of the model to be estimated. Parameters are simply described as a vector of the column names in the `data.frame` provided to the `data` argument.

The `logitr()` function assumes that the deterministic part of the utility function is linear in parameters, i.e.:

```{r, child=here::here('vignettes', 'includes', '_utilityPreferenceObserved.Rmd')}
```

Accordingly, each parameter in the `parNames` argument is an additive part of $v_{j}$. For example, if the observed utility was $v_{j} = \beta x_{j} - \alpha p_{j}$, where $p_{j}$ is price and $x_{j}$ is size, then the `parNames` argument should be `c("price", "size")`, and the `data.frame` used for the `data` argument should have columns called `price` and `size`.

## WTP space models

For models estimated in the WTP space, the `modelSpace` argument should be set to `modelSpace = "wtp"`. The `parNames` argument should **not** contain the name of the column for the price attribute. This is provided separately with the `priceName` argument. For example, if the observed utility was $v_{j} = \lambda \left(\omega' x_{j} - p_{j}\right)$, where $p_{j}$ is price and $x_{j}$ is size, then the `parNames` argument should be `"size"`, and the `priceName` argument should be `"price"`.

Since WTP space models are non-convex, it is recommended that you use a multi-start search to run the optimization loop multiple times (controlled by `numMultiStarts` in the `options` argument) to search for different local minima.

## Mixed logit models

To estimate a mixed logit model, use the `randPars` argument to denote which parameters will be modeled with a distribution. The current package version supports normal and log-normal distributions. For example, if you wanted to model a covariate called `"size"` with a normal distribution, set `randPars = c(size = "n")`. Likewise, if you wanted it to be log-normally distributed, set `randPars = c(size = "ln")`.

Since mixed logit models are non-convex, it is recommended that you use a multi-start search to run the optimization loop multiple times (controlled by `numMultiStarts` in the `options` argument) to search for different local minima. Note that mixed logit models can take a long time to estimate, so setting large number for `numMultiStarts` could take hours or more to complete.

## View results

Use the `summary()` function to print a summary of the results from an estimated model. The function will print the following based on the model settings:

- For a single model run, it prints some summary information, including the model space (Preference or WTP), log-likelihood value at the solution, and a summary table of the model coefficients.
- For MXL models, the function also prints a summary of the random parameters.
- If you set `keepAllRuns = TRUE` in the `options` argument, `summary()` will print a summary of all the multistart runs followed by a summary of the best model (as determined by the largest log-likelihood value).

Use `statusCodes()` to print a description of each status code from the `nloptr` optimization routine.

## Computing and comparing WTP

For models in the preference space, you can get a summary table of the implied WTP by using:

```
wtp(model_pref, priceName)
```

To compare the WTP between two equivalent models in the preference space and WTP spaces, use:

```
wtpCompare(model_pref, model_wtp, priceName)
```

## Simulate shares

Once you have estimated a model, you can use it to simulate the expected shares of a particular set of alternatives. This can be done using the function `simulateShares()`. The simulation reports the expected share as well as a confidence interval for each alternative:

```{r, eval=FALSE}
shares = simulateShares(model, alts, priceName = NULL, alpha = 0.025)
```
