---
title: "Basic Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Usage}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.retina = 3,
  comment = "#>"
)
```

```{r, child='../man/rmdchunks/header.Rmd'}
```

## Data format

```{r, child='../man/rmdchunks/dataFormat.Rmd'}
```

The ["Data Formatting and Encoding"](data_formatting.html) vignette has more details about the required data format.

## Model specification interface

Models are specified and estimated using the `logitr()` function. The `data` argument should be set to the data frame containing the data, and the `outcome` and `obsID` arguments should be set to the column names in the data frame that correspond to the dummy-coded outcome (choice) variable and the observation ID variable, respectively. All variables to be used as model covariates should be provided as a vector of column names to the `pars` argument. Each variable in the vector is additively included as a covariate in the utility model, with the interpretation that they represent utilities in preference space models and WTPs in a WTP space model.

For example, consider a preference space model where the utility for yogurt is given by the following utility model:

```{r, child='../man/rmdchunks/yogurtUtilityPref.Rmd'}
```

\noindent
where $p_{j}$ is `price`, $x_{j1}$ is `feat`, and $x_{j2-4}$ are dummy-coded variables for each `brand` (with the fourth brand representing the reference level). This model can be estimated using the `logitr()` function as follows:

```{r}
library("logitr")

mnl_pref <- logitr(
    data    = yogurt,
    outcome = "choice",
    obsID   = "obsID",
    pars    = c("price", "feat", "brand")
)
```

The equivalent model in the WTP space is given by the following utility model:

```{r, child='../man/rmdchunks/yogurtUtilityWtp.Rmd'}
```

\noindent
To specify this model, simply move `"price"` from the `pars` argument to the `scalePar` argument:

```{r}
mnl_wtp <- logitr(
    data    = yogurt,
    outcome = "choice",
    obsID   = "obsID",
    pars    = c("feat", "brand"),
    scalePar = "price"
)
```

\noindent
In the above model, the variables in `pars` are marginal WTPs, whereas in the preference space model they are marginal utilities. Note that in the WTP space model price is separately specified with the `scalePar` argument as it acts as a scaling term and does not reflect a marginal WTP.

Interactions between covariates can be entered in the `pars` vector separated by the `*` symbol. For example, an interaction between `price` with `feat` in the above preference space model could be included by specifying `pars = c("price", "feat", "brand", "price*feat")`, or even more concisely just `pars = c("price*feat", "brand")` as the interaction between `price` and `feat` will produce individual parameters for `price` and `feat` in addition to the interaction parameter.

Both of these examples are multinomial logit models with fixed parameters. See the ["Estimating Multinomial Logit Models"](mnl_models.html) vignette for more details.

## Parallelized multi-start estimation

Since WTP space models have non-convex log-likelihood functions, it is recommended to use a multi-start search to run the optimization loop multiple times to search for different local minima. This is implemented using the `numMultiStarts` argument, e.g.:

```{r}
mnl_wtp <- logitr(
    data    = yogurt,
    outcome = "choice",
    obsID   = "obsID",
    pars    = c("feat", "brand"),
    scalePar = "price",
    numMultiStarts = 10
)
```

The multi-start is parallelized by default for faster estimation, and the number of cores / workers to use can be manually set using the `numCores` argument. If `numCores` is not provide, then the number of cores is set to `parallel::detectCores() - 1` (the max cores allowed is capped at `parallel::detectCores()`).

## Viewing results

Use the `summary()` function to print a summary of the results from an estimated model, e.g.

```{r}
summary(mnl_pref)
```

Use `statusCodes()` to print a description of each status code from the `nloptr` optimization routine.

You can also extract other values of interest at the solution, such as:

**The estimated coefficients**

```{r}
coef(mnl_pref)
```

**The coefficient standard errors**

```{r}
se(mnl_pref)
```

**The log-likelihood**

```{r}
logLik(mnl_pref)
```

**The variance-covariance matrix**

```{r}
vcov(mnl_pref)
```

## Computing and comparing WTP

For models in the preference space, a summary table of the computed WTP based on the estimated preference space parameters can be obtained with the `wtp()` function. For example, the computed WTP from the previously estimated fixed parameter model can be obtained with the following command:

```{r}
wtp(mnl_pref, scalePar = "price")
```

The `wtp()` function divides the non-price parameters by the negative of the scalePar parameter (price). Standard errors are estimated using the Krinsky and Robb parametric bootstrapping method [@Krinsky1986]. Similarly, the `wtpCompare()` function can be used to compare the WTP from a WTP space model with that computed from an equivalent preference space model:

```{r}
wtpCompare(mnl_pref, mnl_wtp, scalePar = "price")
```

## Mixed logit models

To estimate a mixed logit model, use the `randPars` argument in the `logitr()` function to denote which parameters will be modeled with a distribution. The current package version supports normal (`"n"`) and log-normal (`"ln"`) distributions.

For example, assume the observed utility for yogurts was $v_{j} = \alpha p_{j} + \beta_1 x_{j1} + \beta_2 x_{j2} + \beta_3 x_{j3} + \beta_4 x_{j4}$, where $p_{j}$ is `price`, $x_{j1}$ is `feat`, and $x_{j2-4}$ are dummy-coded variables for `brand`. To model `feat` as well as each of the brands as normally-distributed, set `randPars = c(feat = "n", brand = "n")`:

```{r, eval=FALSE}
mxl_pref <- logitr(
    data     = yogurt,
    outcome  = 'choice',
    obsID    = 'obsID',
    pars     = c('price', 'feat', 'brand'),
    randPars = c(feat = 'n', brand = 'n'),
    numMultiStarts = 10
)
```

Since mixed logit models also have non-convex log-likelihood functions, it is again recommended to use a multi-start search to run the optimization loop multiple times to search for different local minima.

See the ["Estimating Mixed Logit Models"](mxl_models.html) vignette for more details.

## Predicting probabilities and outcomes

Estimated models can be used to predict probabilities and outcomes for a set (or multiple sets) of alternatives based on an estimated model. As an example, consider predicting probabilities for two of the choice observations from the `yogurt` dataset:

```{r}
data <- subset(
  yogurt, obsID %in% c(42, 13),
  select = c('obsID', 'alt', 'choice', 'price', 'feat', 'brand')
)

data
```

In the example below, the probabilities for these two sets of alternatives are computed using the fixed parameter `mnl_pref` model using the `predict()` method:

```{r}
probs <- predict(
  mnl_pref,
  newdata = data,
  obsID = "obsID", 
  ci = 0.95
)

probs
```

The resulting `probs` data frame contains the expected probabilities for each alternative. The lower and upper predictions reflect a 95% confidence interval (controlled by the `ci` argument), which are estimated using the Krinsky and Robb parametric bootstrapping method [@Krinsky1986]. The default is `ci = NULL`, in which case no CI predictions are made.

WTP space models can also be used to predict probabilities, but the `scalePar` argument must be specified for predictions being made on a new data set:

```{r}
probs <- predict(
  mnl_wtp,
  newdata  = data,
  obsID    = "obsID",
  scalePar = "price"
)

probs
```

You can also use the `predict()` method to predict outcomes by setting `type = "outcome"` (the default value is `"prob"` for predicting probabilities). If no new data are provided for the `newdata` argument, then outcomes will be predicted for every alternative in the original data used to estimate the model. In the example below the `returnData` argument is also set to `TRUE` so that the predicted outcomes can be compared to the actual ones.

```{r}
outcomes <- predict(
  mnl_pref,
  type = "outcome",
  returnData = TRUE
)

head(outcomes[c('obsID', 'choice', 'predicted_outcome')])
```

You can quickly compute the accuracy by dividing the number of correctly predicted choices by the total number of choices:

```{r}
chosen <- subset(outcomes, choice == 1)
chosen$correct <- chosen$choice == chosen$predicted_outcome
sum(chosen$correct) / nrow(chosen)
```

See the ["Predicting Probabilities and Choices from Estimated Models"](predict.html) vignette for more details about making predictions.
