---
title: "Data Formatting and Encoding"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Formatting and Encoding}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.retina = 3,
  comment = "#>"
)
library(logitr)
# Read in results from already estimated models  so that the
# examples aren't actually run when building this page, otherwise it'll
# take much longer to build
model_default <- readRDS(
  here::here('inst', 'extdata', 'encoding_model_default.Rds'))
model_character_price <- readRDS(
  here::here('inst', 'extdata', 'encoding_model_character_price.Rds'))
model_dummy_price <- readRDS(
  here::here('inst', 'extdata', 'encoding_model_dummy_price.Rds'))
```

```{r, child=here::here('man', 'rmdchunks', 'header.Rmd')}
```

# Basic required format

```{r, child=here::here('man', 'rmdchunks', 'dataFormat.Rmd')}
```

# Continuous and discrete variables

**Numeric variables** are by default estimated with a single "slope" coefficient.

_Example_: Consider a data frame that contains a `price` variable with the following three levels: `c(10, 15, 20)`. Adding `price` to the `parNames` argument in the main `logitr()` function would result in a single `price` coefficient for the "slope" of the change in price.

**Categorical variables** (i.e. "character" or "factor" type variables) are by default estimated with a coefficient for all but the first "level", which serves as the "baseline" or `"0"` level. Categorical variables are automatically "dummy" coded: `0` for `FALSE` and `1` for `TRUE`.

_Example_: Consider a data frame that contains a `brand` variable with the following four levels: `c("dannon", "hiland", "weight", "yoplait")`. Adding `brand` to the `parNames` argument in the main `logitr()` function would result in three covariates: `brand_hiland`, `brand_weight`, and `brand_yoplait`, with `brand_dannon` serving as the "dummied out" baseline level.

# Creating dummy coded variables

To model a continuous variable as a discrete variable with a coefficient for all but the first level, there are two options:

1. Convert the variable to a `"character"` or `"factor"` type.
2. Create dummy coded variables using the `dummyCode()` function.

The second approach of using the `dummyCode()` function allows the modeler to specify the baseline level. It can also be used to create dummy-coded variables of a categorical variable.

Details for each approach are provided below.

## 1) Convert variable types

The simplest way to model a continuous variable as a discrete variable is to convert the column in the data frame to a `"character"` or `"factor"` type prior to estimating the model. For example, consider the following model:

```{r, eval=FALSE}
model_default <- logitr(
  data       = cars_us,
  choiceName = 'choice',
  obsIDName  = 'obsnum',
  parNames   = c(
    'price', 'hev', 'phev10', 'phev20', 'phev40', 'bev75', 'bev100',
    'bev150', 'american', 'japanese', 'chinese', 'skorean',
    'phevFastcharge', 'bevFastcharge','opCost', 'accelTime')
)
```
```
#> Running Model...
#> Done!
```

In this model, since the `price` variable is a `"double"` variable type, it is by default modeled as a continuous variable with a single "slope" coefficient:

```{r}
typeof(cars_us$price)
summary(model_default)
```

To model `price` as a categorical variable, simple change it to a `"character"` or `"factor"` type:

```{r}
cars_us$price <- as.character(cars_us$price)
typeof(cars_us$price)
```

Now re-estimate the model:

```{r, eval=FALSE}
model_character_price <- logitr(
  data       = cars_us,
  choiceName = 'choice',
  obsIDName  = 'obsnum',
  parNames   = c(
    'price', 'hev', 'phev10', 'phev20', 'phev40', 'bev75', 'bev100',
    'bev150', 'american', 'japanese', 'chinese', 'skorean',
    'phevFastcharge', 'bevFastcharge','opCost', 'accelTime')
)
```
```
#> Running Model...
#> Done!
```

Now `price` is modeled as a categorical variable with a coefficient for all but the first level:

```{r}
typeof(cars_us$price)
summary(model_character_price)
```

## 2) Create dummy-coded variables

For the second option, you can use the `dummyCode()` function to create new dummy-coded variables for all the levels of a continuous variable and then use those variables in the model:

```{r}
cars_us_dummy <- dummyCode(df = cars_us, vars = "price")
names(cars_us_dummy)
```

The new `cars_us_dummy` data frame now contains variables for each level of the `price` column. This approach allows the modeler to specify the baseline level. In this example, I'll use the `price_50` level as the baseline:

```{r, eval=FALSE}
model_dummy_price <- logitr(
  data       = cars_us_dummy,
  choiceName = 'choice',
  obsIDName  = 'obsnum',
  parNames   = c(
    "price_15", "price_18", "price_23", "price_32",
    'hev', 'phev10', 'phev20', 'phev40', 'bev75', 'bev100',
    'bev150', 'american', 'japanese', 'chinese', 'skorean',
    'phevFastcharge', 'bevFastcharge','opCost', 'accelTime')
)
```
```
#> Running Model...
#> Done!
```

Now `price` is modeled with a specified coefficient for all but the `price_50` level:

```{r}
summary(model_dummy_price)
```
