---
title: "Data Formatting and Encoding"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Formatting and Encoding}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, child=here::here('man', 'rmdchunks', 'header.Rmd')}
```

# Basic required format

```{r, child=here::here('man', 'rmdchunks', 'dataFormat.Rmd')}
```

# Continuous and discrete variables

## Default settings

Numeric variables by default are estimated with a single covariate. For example, consider a data frame that contains a `price` variable with three levels `c(10, 15, 20)`. Adding `price` to the `parNames` argument in the main `logitr()` function would estimate a single `price` covariate for the "slope" of the change in price between the values of `10` and `20`.

Categorical variables (i.e. "character" or "factor" type variables) are estimated with a single covariate for every unique "level" except for the first level (the baseline or `"0"` level) using "dummy" encoding (`0` for `FALSE` and `1` for `TRUE`). For example, consider a data frame that contains a `brand` variable with four levels `c("dannon", "hiland", "weight", "yoplait")`. Adding `brand` to the `parNames` argument in the main `logitr()` function would estimate three covariates: `brand_hiland`, `brand_weight`, and `brand_yoplait`, with `brand_dannon` serving as the "dummied out" baseline level.

## Dummy coding variables

To model a continuous variable as a discrete variable with a covariate for all but the first "level", there are two options:

1. Convert the variable to a "character" or "factor" type.
2. Create dummy coded variables using the `dummyCode()` function.
