---
title: "Estimating Models with Interactions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimating Models with Interactions}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.retina = 3,
  comment = "#>"
)
library(logitr)
# Read in results from already estimated models  so that the
# examples aren't actually run when building this page, otherwise it'll
# take much longer to build
model_price_feat <- readRDS(
  here::here('inst', 'extdata', 'int_model_price_feat.Rds'))
model_price_brand <- readRDS(
  here::here('inst', 'extdata', 'int_model_price_brand.Rds'))
model_price_feat_mxl <- readRDS(
  here::here('inst', 'extdata', 'model_price_feat_mxl.Rds'))
```

# Interactions with continuous variables

To add interactions between coefficients in your model, you can add additional arguments in the `parNames` vector in the `logitr()` function separated by the `*` symbol. For example, let's say we want to interact `price` with `feat` in the following model:

```{r, eval=FALSE}
model <- logitr(
  data       = yogurt,
  choiceName = 'choice',
  obsIDName  = 'obsID',
  parNames   = c('price', 'feat', 'brand')
)
```

To do so, I could add `"price*feat"` to the `parNames` vector:

```{r, eval=FALSE}
model_price_feat <- logitr(
  data       = yogurt,
  choiceName = 'choice',
  obsIDName  = 'obsID',
  parNames   = c('price', 'feat', 'brand', 'price*feat')
)
```
```
#> Running Model...
#> Done!
```

The model now has an estimated coefficient for the `price*feat` effect:

```{r}
summary(model_price_feat)
```

# Interactions with discrete variables

In the above example, both `price` and `feat` were continuous variables, so only a single interaction coefficient was needed.

In the case of interacting _discrete_ variables, multiple interactions coefficients will be estimated according to the number of levels in the discrete attribute. For example, the interaction of `price` with `brand` will require three new interactions - one for each level of the `brand` variable except the first "baseline" level:

```{r, eval=FALSE}
model_price_brand <- logitr(
  data       = yogurt,
  choiceName = 'choice',
  obsIDName  = 'obsID',
  parNames   = c('price', 'feat', 'brand', 'price*brand')
)
```
```
#> Running Model...
#> Done!
```

The model now has three estimated coefficients for the `price*brand` effect:

```{r}
summary(model_price_brand)
```

# Interactions in mixed logit models

Suppose I want to include an interaction between two variables and I also want one of those variables to be modeled as normally distributed across the population. The example below illustrates this cases, where a `price*feat` interaction is specified and the `feat` parameter is modeled as normally distributed by setting `randPars = c(feat = "n")`:randPars   = c(feat = "n")

```{r, eval=FALSE}
model_price_feat_mxl <- logitr(
  data       = yogurt,
  choiceName = 'choice',
  obsIDName  = 'obsID',
  parNames   = c('price', 'feat', 'brand', 'price*feat'),
  randPars   = c(feat = "n")
)
```
```
#> Running Model...
#> Done!
```

In this case, the `price*feat` interaction parameter is interpreted as a difference in the `feat_mu` parameter and price; that is, it an interaction in the _mean_ `feat` parameter and `price`:

```{r}
summary(model_price_brand)
```

