---
title: "Estimating mixed logit models"
output:
  html_document:
    number_sections: false
    toc: no
bibliography: library.bib
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Read in results from already estimated models  so that the
# examples aren't actually run when building this page, otherwise it'll
# take much longer to build
mxl_pref <- readRDS(here::here('vignettes', 'results', 'mxl_pref.Rds'))
mxl_wtp  <- readRDS(here::here('vignettes', 'results', 'mxl_wtp.Rds'))
wtp_mxl_pref <- readRDS(here::here('vignettes', 'results', 'wtp_mxl_pref.Rds'))
wtp_mxl_comparison <- readRDS(here::here('vignettes', 'results', 'wtp_mxl_comparison.Rds'))
```

# The data

```{r, child=here::here('vignettes', 'includes', '_yogurtDataDescription.Rmd')}
```

# Preference space model

This example will estimate the following mixed logit model in the preference space:

```{r, child=here::here('vignettes', 'includes', '_mxlPrefExample.Rmd')}
```

where the parameters $\alpha$, $\beta_1$, $\beta_2$, $\beta_3$, and $\beta_4$ have units of utility, and the parameter for $x_{j}^{\mathrm{FEAT}}$ is normally distributed.

To estimate the model, first load the **logitr** package:

```{r eval=FALSE}
library(logitr)
```

Estimate the model using the `logitr()` function:

```{r eval=FALSE}
mxl_pref <- logitr(
    data       = yogurt,
    choiceName = "choice",
    obsIDName  = "obsID",
    parNames   = c("price", "feat", "dannon", "hiland", "yoplait"),
    randPars   = c(feat = "n"),
    options    = list(
        # You should run a multistart for MXL models since they are non-convex,
        # but it can take a long time. Here I just use 1 for brevity:
        numMultiStarts = 1,
        numDraws       = 500))
```
```
Running Model
Done!
```

Print a summary of the results:

```{r, eval=FALSE}
summary(mxl_pref)
```
```
=================================================
MODEL SUMMARY: 
                         
Model Space:   Preference
Model Run:         1 of 1
Iterations:            36
Elapsed Time:   0h:2m:39s
Weights Used?:      FALSE

Model Coefficients: 
            Estimate StdError    tStat   pVal signif
price      -0.392768 0.026708 -14.7061 0.0000    ***
feat.mu     0.351933 0.204608   1.7200 0.0856      .
dannon      0.663863 0.055794  11.8984 0.0000    ***
hiland     -3.324270 0.164101 -20.2575 0.0000    ***
yoplait     1.458054 0.095512  15.2656 0.0000    ***
feat.sigma  2.360350 0.515567   4.5782 0.0000    ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Model Fit Values: 
                                     
Log.Likelihood.         -2645.9202707
Null.Log.Likelihood.    -3343.7419990
AIC.                     5303.8405000
BIC.                     5338.5698000
McFadden.R2.                0.2086948
Adj..McFadden.R2            0.2069005
Number.of.Observations.  2412.0000000

Summary of 10k Draws for Random Coefficients: 
       Min.   1st Qu.    Median      Mean  3rd Qu.     Max.
1 -8.822073 -1.240571 0.3511815 0.3499323 1.942267 8.875534
```

View the estimated model coefficients:

```{r}
coef(mxl_pref)
```

Compute the WTP implied from the preference space model:

```{r, eval=FALSE}
wtp_mxl_pref <- wtp(mxl_pref, priceName = "price")
```
```{r}
wtp_mxl_pref
```

# WTP space model

This example will estimate the following mixed logit model in the WTP space:

```{r, child=here::here('vignettes', 'includes', '_mxlWtpExample.Rmd')}
```

where the parameters $\omega_1$, $\omega_2$, $\omega_3$, and $\omega_4$ have units of dollars and $\lambda$ is the scale parameter, and the WTP parameter for $x_{j}^{\mathrm{FEAT}}$ is normally distributed.

To estimate the model, first load the **logitr** package:

```{r eval=FALSE}
library(logitr)
```

Estimate the model using the `logitr()` function:

```{r eval=FALSE}
# Extract the WTP computed from the preference space model
# to use as the initial starting values
startingValues <- wtp_mxl_pref$Estimate

mxl_wtp <- logitr(
    data       = yogurt,
    choiceName = "choice",
    obsIDName  = "obsID",
    parNames   = c("feat", "dannon", "hiland", "yoplait"),
    priceName  = "price",
    randPars   = c(feat="n"),
    modelSpace = "wtp",
    options = list(
        # You should run a multistart for MXL models since they are non-convex,
        # but it can take a long time. Here I just use 1 for brevity:
        numMultiStarts = 1,
        startVals   = startingValues,
        # Because the computed WTP from the preference space model has values
        # as large as 8, I increase the boundaries of the random starting values:
        startParBounds = c(-5, 5),
        numDraws       = 500))
```
```
Running Model
**Using User Provided Starting Values For This Run**
Done!
```

Print a summary of the results:

```{r, eval=FALSE}
summary(mxl_wtp)
```
```
=================================================
MODEL SUMMARY: 
                                 
Model Space:   Willingness-to-Pay
Model Run:                 1 of 1
Iterations:                   112
Elapsed Time:           0h:9m:41s
Weights Used?:              FALSE

Model Coefficients: 
            Estimate StdError    tStat   pVal signif
lambda      0.392768 0.026708  14.7062 0.0000    ***
feat.mu     0.896032 0.534239   1.6772 0.0936      .
dannon      1.690215 0.172674   9.7885 0.0000    ***
hiland     -8.463701 0.517775 -16.3463 0.0000    ***
yoplait     3.712255 0.161357  23.0065 0.0000    ***
feat.sigma  6.009556 1.323045   4.5422 0.0000    ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Model Fit Values: 
                                     
Log.Likelihood.         -2645.9202707
Null.Log.Likelihood.    -3343.7419990
AIC.                     5303.8405000
BIC.                     5338.5698000
McFadden.R2.                0.2086948
Adj..McFadden.R2            0.2069005
Number.of.Observations.  2412.0000000

Summary of 10k Draws for Random Coefficients: 
       Min.   1st Qu.    Median      Mean  3rd Qu.    Max.
1 -22.46139 -3.158553 0.8941192 0.8909386 4.945094 22.5975
```

View the estimated model coefficients:

```{r}
coef(mxl_wtp)
```

If you want to compare the WTP from the two different model spaces, use the `wtpCompare()` function:

```{r, eval=FALSE}
wtp_mxl_comparison <- wtpCompare(mxl_pref, mxl_wtp, priceName = 'price')
```
```{r}
wtp_mxl_comparison
```

Note that the WTP will not be the same between preference space and WTP space MXL models. This is because the distributional assumptions in MXL models imply different distributions on WTP depending on the model space. See Train and Weeks [-@Train2005] and Sonnier, Ainslie, and Otter [-@Sonnier2007] for details on this topic.

```{r, eval=FALSE, include=FALSE}
# Save results - do this only after manually
saveRDS(mxl_pref,
        here::here('vignettes', 'results', 'mxl_pref.Rds'))
saveRDS(mxl_wtp,
        here::here('vignettes', 'results', 'mxl_wtp.Rds'))
saveRDS(wtp_mxl_pref,
        here::here('vignettes', 'results', 'wtp_mxl_pref.Rds'))
saveRDS(wtp_mxl_comparison,
        here::here('vignettes', 'results', 'wtp_mxl_comparison.Rds'))
```

# References
