---
title: "Mixed logit models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mixed logit models}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
bibliography: "`r here::here('vignettes', 'library.bib')`"
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.retina = 3,
  comment = "#>"
)
library(logitr)
# Read in results from already estimated models  so that the
# examples aren't actually run when building this page, otherwise it'll
# take much longer to build
mxl_pref <- readRDS(here::here('inst', 'extdata', 'mxl_pref.Rds'))
mxl_wtp  <- readRDS(here::here('inst', 'extdata', 'mxl_wtp.Rds'))
wtp_mxl_pref <- readRDS(here::here('inst', 'extdata', 'wtp_mxl_pref.Rds'))
wtp_mxl_comparison <- readRDS(
    here::here('inst', 'extdata', 'wtp_mxl_comparison.Rds'))
```

# The data

```{r, child=here::here('man', 'rmdchunks', 'yogurtDataDescription.Rmd')}
```

# Preference space model

This example will estimate the following mixed logit model in the preference space:

```{r, child=here::here('man', 'rmdchunks', 'mxlPrefExample.Rmd')}
```

where the parameters $\alpha$, $\beta_1$, $\beta_2$, $\beta_3$, and $\beta_4$ have units of utility, and the parameter for $x_{j}^{\mathrm{FEAT}}$ is normally distributed.

To estimate the model, first load the **logitr** package:

```{r eval=FALSE}
library(logitr)
```

Estimate the model using the `logitr()` function:

```{r eval=FALSE}
mxl_pref <- logitr(
  data       = yogurt,
  choiceName = 'choice',
  obsIDName  = 'obsID',
  parNames   = c('price', 'feat', 'hiland', 'yoplait', 'dannon'),
  randPars   = c(feat = 'n', hiland = 'n', yoplait = 'n', dannon = 'n'),
  # You should run a multistart for MXL models since they are non-convex,
  # but it can take a long time. Here I just use 5 starts for brevity:
  options    = list(numMultiStarts = 5)
)
```
```
#> Running Multistart 1 of 5
#> Running Multistart 2 of 5
#> Running Multistart 3 of 5
#> Running Multistart 4 of 5
#> Running Multistart 5 of 5
#> Done!
```

Print a summary of the results:

```{r}
summary(mxl_pref)
```

View the estimated model coefficients:

```{r}
coef(mxl_pref)
```

Compute the WTP implied from the preference space model:

```{r, eval=FALSE}
wtp_mxl_pref <- wtp(mxl_pref, priceName = "price")
```
```{r}
wtp_mxl_pref
```

# WTP space model

This example will estimate the following mixed logit model in the WTP space:

```{r, child=here::here('man', 'rmdchunks', 'mxlWtpExample.Rmd')}
```

where the parameters $\omega_1$, $\omega_2$, $\omega_3$, and $\omega_4$ have units of dollars and $\lambda$ is the scale parameter, and the WTP parameter for $x_{j}^{\mathrm{FEAT}}$ is normally distributed.

To estimate the model, first load the **logitr** package:

```{r eval=FALSE}
library(logitr)
```

Estimate the model using the `logitr()` function:

```{r eval=FALSE}
# Extract the WTP computed from the preference space model
# to use as the initial starting values
startingValues <- wtp_mxl_pref$Estimate

mxl_wtp <- logitr(
  data       = yogurt,
  choiceName = 'choice',
  obsIDName  = 'obsID',
  parNames   = c('feat', 'hiland', 'yoplait', 'dannon'),
  priceName  = 'price',
  randPars   = c(feat = 'n', hiland = 'n', yoplait = 'n', dannon = 'n'),
  modelSpace = 'wtp',
  options    = list(
    # You should run a multistart for MXL models since they are non-convex,
    # but it can take a long time. Here I just use 5 starts for brevity:
    numMultiStarts = 5,
    # Use the computed WTP from the preference space model as the starting
    # values for the first run:
    startVals = startingValues,
    # Because the computed WTP from the preference space model has values
    # as large as 8, I increase the boundaries of the random starting values:
    startParBounds = c(-10, 10)))
```
```
#> Running Multistart 1 of 5
#> **Using User Provided Starting Values For This Run**
#> Running Multistart 2 of 5
#> Running Multistart 3 of 5
#> Running Multistart 4 of 5
#> Running Multistart 5 of 5
#> Done!
```

Print a summary of the results:

```{r}
summary(mxl_wtp)
```

View the estimated model coefficients:

```{r}
coef(mxl_wtp)
```

If you want to compare the WTP from the two different model spaces, use the `wtpCompare()` function:

```{r, eval=FALSE}
wtp_mxl_comparison <- wtpCompare(mxl_pref, mxl_wtp, priceName = 'price')
```
```{r}
wtp_mxl_comparison
```

Note that the WTP will not be the same between preference space and WTP space MXL models. This is because the distributional assumptions in MXL models imply different distributions on WTP depending on the model space. See Train and Weeks [-@Train2005] and Sonnier, Ainslie, and Otter [-@Sonnier2007] for details on this topic.

# References
