---
title: "Predicting Choices from Estimated Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Predicting Choices from Estimated Models}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
bibliography: "`r here::here('vignettes', 'library.bib')`"
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.path   = "figs/",
  fig.retina = 3,
  comment = "#>"
)
library(logitr)
# Read in results from already estimated models  so that the
# examples aren't actually run when building this page, otherwise it'll
# take much longer to build
choices_mnl_pref <- readRDS(
  here::here('inst', 'extdata', 'choices_mnl_pref.Rds'))
choices_mnl_wtp <- readRDS(
  here::here('inst', 'extdata', 'choices_mnl_wtp.Rds'))
choices_mxl_pref <- readRDS(
  here::here('inst', 'extdata', 'choices_mxl_pref.Rds'))
choices_mxl_wtp <- readRDS(
  here::here('inst', 'extdata', 'choices_mxl_wtp.Rds'))
```

Once a model has been estimated, it can be used to predict choices for a set of alternatives. This vignette demonstrates examples of how to so using the `predictChoices()` function along with the results of an estimated model.

# The data

To predict choices, you first need to define a set of alternatives for which you want to make predictions. Each row should be an alternative, and each column should be an attribute. I will predict choices on the full `yogurt` data set, which was used to estimate each of the models used in this example.

```{r, child=here::here('man', 'rmdchunks', 'yogurtDataDescription.Rmd')}
```

```{r}
head(yogurt)
```

# Predicting with multinomial logit models

## Preference space parameterization

In the example below, I estimate a preference space MNL model called `mnl_pref`. I can then use the `predictChoices()` function with the `mnl_pref` model to predict the choices for each set of alternatives in the `yogurt` data set:

```{r, eval=FALSE}
# Estimate the model
mnl_pref <- logitr(
  data       = yogurt,
  choiceName = 'choice',
  obsIDName  = 'obsID',
  parNames   = c('price', 'feat', 'brand')
)

# Make predictions
choices_mnl_pref <- predictChoices(
  model      = mnl_pref,
  alts       = yogurt,
  choiceName = "choice",
  obsIDName  = "obsID"
)
```
```{r}
head(choices_mnl_pref)
```

The results are a data frame showing the actual choices, `choice`, and the predicted choices, `choice_predict`. You can quickly compute the accuracy by dividing the number of correctly predicted choices by the total number of choices:

```{r}
library(dplyr)

choices_mnl_pref %>%
  filter(choice == 1) %>%
  summarise(p_correct = sum(choice_predict == choice) / n())
```

## WTP space parameterization

You can also use WTP space models to predict choices. For example, here are the results from an equivalent model but in the WTP space:

```{r, eval=FALSE}
# Estimate the model
mnl_wtp <- logitr(
  data       = yogurt,
  choiceName = 'choice',
  obsIDName  = 'obsID',
  parNames   = c('feat', 'brand'),
  priceName  = 'price',
  modelSpace = 'wtp',
  options = list(numMultiStarts = 10)
)

# Make predictions
choices_mnl_wtp <- predictChoices(
  model      = mnl_wtp,
  alts       = yogurt,
  choiceName = "choice",
  obsIDName  = "obsID"
)
```
```
#> NOTE: Using results from run 8 of 10 multistart runs
#> (the run with the largest log-likelihood value)
```
```{r}
head(choices_mnl_wtp)
```

# Predicting with mixed logit models

## Preference space parameterization

You can also use mixed logit models to predict choices. Heterogeneity is modeled by simulating draws from the population estimates of the estimated model. Here is an example using a preference space mixed logit model:

```{r, eval=FALSE}
# Estimate the model
mxl_pref <- logitr(
  data       = yogurt,
  choiceName = 'choice',
  obsIDName  = 'obsID',
  parNames   = c('price', 'feat', 'brand'),
  randPars   = c(feat = 'n', brand = 'n'),
  options    = list(numMultiStarts = 5)
)

# Make predictions
choices_mxl_pref <- predictChoices(
  model      = mxl_pref,
  alts       = yogurt,
  choiceName = "choice",
  obsIDName  = "obsID"
)
```
```{r}
head(choices_mxl_pref)
```

## WTP space parameterization

Likewise, mixed logit WTP space models can also be used to predict choices:

```{r, eval=FALSE}
# Estimate the model
mxl_wtp <- logitr(
  data       = yogurt,
  choiceName = 'choice',
  obsIDName  = 'obsID',
  parNames   = c('feat', 'brand'),
  priceName  = 'price',
  randPars   = c(feat = 'n', brand = 'n'),
  modelSpace = 'wtp',
  options    = list(numMultiStarts = 5)
)

# Make predictions
choices_mxl_wtp <- predictChoices(
  model      = mxl_wtp,
  alts       = yogurt,
  choiceName = "choice",
  obsIDName  = "obsID"
)
```
```{r}
head(choices_mxl_wtp)
```

# Compare prediction results

```{r}
# Combine models into one data frame
choices <- rbind(
  choices_mnl_pref, choices_mnl_wtp, choices_mxl_pref, choices_mxl_wtp)
choices$model <- c(
  rep("mnl_pref", nrow(choices_mnl_pref)),
  rep("mnl_wtp",  nrow(choices_mnl_wtp)),
  rep("mxl_pref", nrow(choices_mxl_pref)),
  rep("mxl_wtp",  nrow(choices_mxl_wtp)))

# Compute prediction accuracy by model
choices %>%
  filter(choice == 1) %>%
  mutate(predict_correct = (choice_predict == choice)) %>%
  group_by(model) %>%
  summarise(p_correct = sum(predict_correct) / n())
```

The models all perform about the same with ~38% correct predictions. This is significantly better than random predictions, which should be 25%.

# References
