---
title: "Simulating Shares from Estimated Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulating Shares from Estimated Models}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
bibliography: "`r here::here('vignettes', 'library.bib')`"
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.retina = 3,
  comment = "#>"
)
library(logitr)
# Read in results from already estimated models  so that the
# examples aren't actually run when building this page, otherwise it'll
# take much longer to build
sim_mnl_pref <- readRDS(here::here('inst', 'extdata', 'sim_mnl_pref.Rds'))
sim_mnl_wtp <- readRDS(here::here('inst', 'extdata', 'sim_mnl_wtp.Rds'))
sim_mxl_pref <- readRDS(here::here('inst', 'extdata', 'sim_mxl_pref.Rds'))
sim_mxl_wtp <- readRDS(here::here('inst', 'extdata', 'sim_mxl_wtp.Rds'))
```

Once a model has been estimated, it can be used to simulate the predicted shares for a set of alternatives. This vignette demonstrates examples of how to so using the `simulateShares()` function along with the results of an estimated model.

# The data

```{r, child=here::here('man', 'rmdchunks', 'yogurtDataDescription.Rmd')}
```

# Simulating shares from estimated models

 To simulate shares, you first need to create a set of alternatives where each row is an alternative and each column an attribute. In this example, I'll just use one of the choice observations from the `yogurt` dataset:

```{r}
alts <- subset(yogurt, obsID == 42,
               select = c('feat', 'price', 'hiland', 'weight', 'yoplait'))
row.names(alts) <- c('dannon', 'hiland', 'weight', 'yoplait')
alts
```

Let's say we have estimated a preference space MNL model called `mnl_pref`. We can use the `simulateShares()` function with the `mnl_pref` model to predict the shares for our `alts` set of alternatives:

```{r, eval=FALSE}
sim_mnl_pref <- simulateShares(mnl_pref, alts, alpha = 0.025)
```
```{r}
sim_mnl_pref
```

The results show the expected shares for each alternative. The low and high values show a 95% confidence interval, estimated using simulation. You can change the CI level by setting alpha to a different value (e.g. a 90% CI is obtained with alpha = 0.05).

You can also use WTP space models to simulate shares, but you must provide the additional `priceName` argument to the `simulateShares()` function. For example, here are the results from an equivalent model but in the WTP space:

```{r, eval=FALSE}
sim_mnl_wtp <- simulateShares(mnl_wtp, alts, priceName = 'price')
```
```{r}
sim_mnl_wtp
```

Since these two models are equivalent except in different spaces, the simulation results should be the same.

Simulations also work with mixed logit models, and they account for how heterogeneity is modeled by simulating draws from the population estimates of the estimated model:

```{r, eval=FALSE}
sim_mxl_pref <- simulateShares(mxl_pref, alts)
```
```{r}
sim_mxl_pref
```

Likewise, mixed logit WTP space models can also be used to simulate shares:

```{r, eval=FALSE}
sim_mxl_wtp <- simulateShares(mxl_wtp, alts, priceName = 'price')
```
```{r}
sim_mxl_wtp
```
