---
title: "Summarizing Results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Usage}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.retina = 3,
  comment = "#>"
)
```

```{r, child='../man/rmdchunks/header.Rmd'}
```

# Extracting summary tables 

The most common function used to view an overall summary of a model is the `summary()` function:

```{r}
library(logitr)

model <- logitr(
  data    = yogurt,
  outcome = "choice",
  obsID   = "obsID",
  pars    = c("price", "feat", "brand")
)

summary(model)
```

The summary prints out a table of the model coefficients as well as other information about the model, such as the log-likelihood, the number of observations, etc.

You can extract the coefficient from the summary table as a `data.frame` using `coef(summary(model))`:

```{r}
coefs <- coef(summary(model))
coefs
```

Another approach is to use the `tidy()` function from the [{broom}](https://broom.tidymodels.org/index.html) package:

```{r}
library(broom)

coefs <- tidy(mnl_pref)
coefs
```

The `tidy()` function returns a `tibble` and provides a more standardized output and interfaces well with other packages. You can also append a confidence interval to the data frame:

```{r}
coefs <- tidy(mnl_pref, conf.int = TRUE)
coefs
```

# Extracting other values 

You can also extract other values of interest at the solution, such as:

**The estimated coefficients**

```{r}
coef(mnl_pref)
```

**The estimated standard errors**

```{r}
se(mnl_pref)
```

**The log-likelihood**

```{r}
logLik(mnl_pref)
```

**The variance-covariance matrix**

```{r}
vcov(mnl_pref)
```

You can also view a summary of statistics about the model using the `glance()` function from the [{broom}](https://broom.tidymodels.org/index.html) package:

```{r}
glance(model)
```

# Formatted summary tables

## gtsummary

Often times you will need to create summary tables that are formatted for publication. The [{gtsummary}(https://www.danieldsjoberg.com/gtsummary/)] package offers a convenient solution that works well with `logitr`models. For example, a formatted summary table can be obtained using the `tbl_regression()` function:

```{r}
library(gtsummary)

model |> 
  tbl_regression() |> 
  as_kable()
```

The `tbl_regression()` function has many [options](https://www.danieldsjoberg.com/gtsummary/articles/tbl_regression.html) for customizing the output table. For example, you can change the coefficient names with the `label` argument:

```{r}
model |> 
  tbl_regression(
    label = list(
        feat = "Newspaper ad shown?",
        brand = "Yogurt's brand"
    )
  ) |> 
  as_kable()
```

## texreg

Another option for obtaining a formatted table is to use the [{texreg}](https://github.com/leifeld/texreg) package. This is particularly useful for obtaining tables formatted for use in [LaTeX](https://www.latex-project.org/).

