
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Read in results from already estimated models, otherwise
# the page building will take forever!
mxl_pref <- readRDS(here::here('examples', 'results', 'mxl_pref.Rds'))
mxl_wtp  <- readRDS(here::here('examples', 'results', 'mxl_wtp.Rds'))
wtp_mxl_comparison <- readRDS(here::here('examples', 'results',
  'wtp_mxl_comparison.Rds'))
```

First, load the **logitr** library:
```{r, message=FALSE, warning=FALSE}
library('logitr')
```

We'll use the Yogurt data set from the [**mlogit** package](https://cran.r-project.org/web/packages/mlogit/index.html)
```{r}
head(yogurt)
```

## Mixed logit models in the preference space

To estimate the model, use the `logitr()` function. In the example below, the `data` argument is set to the name of the data.frame storing the choice observations (in this case, `yogurt`). The `choiceName` and `obsIDName` arguments are the names of the columns in the data.frame that identify the following:

- `choiceName`: A dummy variable that identifies which alternative was chosen (`1` = chosen, `0` = not chosen).
- `obsIDName`: A sequence of numbers that identifies each unique choice occasion. For example, if the first three choice occasions had 2 alternatives each, then the first 9 rows of the `obsID` variable would be `1, 1, 2, 2, 3, 3`.

The `parNames` argument is a vector of the column names in the data.frame that will be used to estimate parameters in the model. Must be the same as the column names in the `data` argument. For WTP space models, do not include price in `parNames`.

The model assumes that the deterministic part of the utility function is linear in parameters (*v* = *beta* ' *x*). Accordingly, each parameter in the `parNames` argument is an additive part of *v*. For example, for the utility model *u* = *beta1* * *price* + *beta2* * *size* + *error*, then the `parNames` argument should be `c('price', 'size')`, and there should be two columns of data in the `data` argument called `price` and `size`. If you wanted to add a third parameter, say *price^2*, then you should create a separate variable in the data.frame called something like `priceSquared` and your `parNames` argument would be `c('price', 'size', 'priceSquared')`.

**WTP space models**:
The `parNames` should be the WTP parameters, and the `price` parameter is denoted by the separate argument `priceName`. For example, for the utility model *u* = *lambda*(*beta1* * *size* - *price*) + *error*, then the `parNames` argument should be `c('size')` and the `priceName` argument should be `'price'`.

`randPars` argument is used to denote which parameters will be modeled with a distribution. In this example, only the `'feat'` attribute is modeled as normally distributed, so all other parameters will be modeled with fixed coefficients.

Since mixed logit models are non-convex, it is recommended that you use a multi-start search to run the optimization loop multiple times (controlled by the `numMultiStarts` argument in `options`) to search for different local minima. In this example, `numMultiStarts` is set to `1` for brevity.

```{r, eval=FALSE}
mxl_pref <- logitr(
  data       = yogurt,
  choiceName = 'choice',
  obsIDName  = 'obsID',
  parNames   = c('price', 'feat', 'dannon', 'hiland', 'yoplait'),
  randPars   = c(feat = 'n'),
  options    = list(
    numMultiStarts = 1,
    numDraws       = 500))
```
```
Running Model
Done!
```

You can view a summary of the model output with the `summary()` function:
```{r}
summary(mxl_pref)
```

Since this model is a "preference" space model, you can use the result to compute the implied willingness to pay with the `wtp()` function:
```{r}
wtp_mxl_pref = wtp(mxl_pref, priceName = 'price')
wtp_mxl_pref
```

## Mixed logit models in the WTP space

To estimate a mixed logit model in the WTP space, you have to add the `modelSpace = "wtp"` argument. You also must provide the `priceName` argument, which specified the name of the column in the `data` argument that is price. In this example, I model price with a fixed coefficient, but you could model it as a random parameter by specifying the `randPrice` argument (typically set to `'ln'` for log-normal).

Note also that I am using the computed WTP from the preference space model as the starting parameters. These values will only be used as the starting guess for the first iteration of a multi-start run.
```{r, eval=FALSE}
mxl_wtp = logitr(
  data       = yogurt,
  choiceName = 'choice',
  obsIDName  = 'obsID',
  parNames   = c('feat', 'dannon', 'hiland', 'yoplait'),
  priceName  = 'price',
  randPars   = c(feat = 'n'),
  # randPrice  = 'ln',
  modelSpace = 'wtp',
  options = list(
    numMultiStarts = 1,
    startVals      = wtp_mxl_pref$Estimate,
    startParBounds = c(-5, 5),
    numDraws       = 500))
```
```
Running Model
Done!
```

You can view a summary of the model output with the `summary()` function:
```{r}
summary(mxl_wtp)
```

If you want to compare the WTP from the two different model spaces, use the `wtpCompare()` function:
```{r, eval=FALSE}
wtp_mxl_comparison <- wtpCompare(mxl_pref, mxl_wtp, priceName = 'price')
```
```{r}
wtp_mxl_comparison
```

Note that the WTP will not be the same between preference space and WTP space MXL models. This is because the distributional assumptions in MXL models imply different distributions on WTP depending on the model space. See Train and Weeks (2005) and Sonnier, Ainslie, and Otter (2007) for details on this topic:

- Train K, Weeks M (2005). “Discrete Choice Models in Preference Space and Willingness- to-Pay Space.” In R Scarpa, A Alberini (eds.), Applications of Simulation Methods in Environmental and Resource Economics, volume 6 of The Economics of Non-Market Goods and Resources, pp. 1-16. Springer-Verlag.
- Sonnier G, Ainslie A, Otter T (2007). "Heterogeneity Distributions of Willingness-to-Pay in Choice Models." Quantitative Marketing and Economics, 5(3), 313–331.

```{r, eval=FALSE}
# Save results
saveRDS(mxl_pref,
        here::here('examples', 'results', 'mxl_pref.Rds'))
saveRDS(mxl_wtp,
        here::here('examples', 'results', 'mxl_wtp.Rds'))
saveRDS(wtp_mxl_comparison,
        here::here('examples', 'results', 'wtp_mxl_comparison.Rds'))
```
